---
title: "Chapter_9_Exercises"
author: "Colby S."
date: "4/8/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---
General Statistics

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

get a  basic statistical summary of your data
```{r 9.1 Summarizing Your Data}
# predict the output of the summary function to each of the variables then apply the summary function. Did the output surprise you? 
# bonus: share a way to get more meaningful output

## VECTOR
vec <- rnorm(100, mean = 100, sd = 15)
# summary(vec)

## FACTOR
factor_lvls <- factor(c("yes", "no", "maybe"), levels = c("yes", "no", "maybe"))
fac <- sample(factor_lvls, size = 100, prob = c(0.6, 0.3, 0.1), replace = TRUE)
# summary(fac)

## CHARACTER VECTOR
char <- c("What is returned when you summarise a character vector with one string")
# summary(char)
# length(char)
char2 <- c("how", "about", "when", "there's", "two", "characters")
# summary(char2)
# bonus: how many characters (letters/spaces/punctuation/etc) are in the vector "char")
nchar(char2)

## MATRIX
mat <- cbind(
  rnorm(100, mean = 100, sd = 15),
  rnorm(100, mean = 50, sd = 30),
  rnorm(100, mean = 80, sd = 5)
)
# summary(mat)

## DATAFRAME
# Apply the summary function to df.
df <- data.frame(
  char, fac, vec
)
# summary(df)

## LIST
# apply the summary function to each element in list1
list1 <- list(2, "hello", c(3,5,4), 1:5, list(FALSE, c("this", "is","a","list"),c(FALSE,TRUE,TRUE,TRUE,FALSE)))
summary(list1)
purrr::map(list1, summary)
```

count the relative frequency of certain observations in your sample
```{r 9.2 Calculating Relative Frequencies}
set.seed(120)
df_effects <- data.frame(
  before = rnorm(1000, 100, 20),
  after = rnorm(1000, 120, 20),
  label = sample(c("NJ", "FL"), 1000, prob = c(0.6, 0.4), replace = TRUE)
)
before <- as.vector(df_effects$before)
after <- as.vector(df_effects$after)
label <- as.vector(df_effects$label)

time_series <- rnorm(1000, 100, 20)
# what fraction of labels are NJ
 mean(label == "FL")
# for what fraction of cases is before greater than after
 mean(before > after)
 # what fraction of before values exceed two standard deviations from the mean of before values
 mean(before - mean(before) > 2*sd(before))
 # fraction of time series values that are larger than the value taken immediately prior
 mean(diff(time_series) > 0)
```

tabulate one factor or build a contingency table from multiple factors
```{r 9.3 Tabulating Factors and Creating Contingency Tables}
f1 <- factor(
  sample(letters[1:5], size = 100, prob = c(17,17,22,20,24), replace = TRUE),
  levels = c("a", "b", "c", "d", "e"))

f2 <- factor(
  sample(letters[6:8], size = 100, replace = TRUE),
  levels = c("f", "g", "h"))

# produce counts of factor f1
table(f1)
# produce a contingency table for f1 and f2
table(f1, f2)
# will the table function work with three factors?

# also see xtabs
```

test two categorical variables for independence using the chi-squared test
```{r 9.4 Testing Categorical Variables for Independence}
set.seed(42)
initial <- factor(sample(c("Yes", "No", "Maybe"), 100, replace = TRUE))
outcome <- factor(sample(c("Pass", "Fail"), 100, replace = TRUE))
summary(table(initial,outcome))
# chisq.test() also does this 
```

Given a fraction f, you want to know the corresponding quantile of your data.
```{r 9.5 Calculating Quantiles (and Quartiles) of a Dataset}
# find the observation at the 95th quantile of var
set.seed(42)
var <- rnorm(100, mean = 100, sd = 20)
quantile(vec, 0.95)
# find the observations expected at each quartile (you should only call one command)
quantile(vec)
# which observation in vec delimits the lower 5% of the data
quantile(vec, 0.05)
# using only one command, return the observations at the 5th and 95th percentile of vec
quantile(vec, c(0.5, 0.95))
# what is the types argument and how many types are there?
```

Given an observation x from your data, what is its corresponding quantile
```{r 9.6 Inverting a Quantile}
# what quantile is a value of 100 in vector "vec"
mean(vec<100)
```

calculate the corresponding z-scores for all data elements
AKA Normalize the data
```{r 9.7 Converting Data to Z-Scores}
# what is a z score
# convert the data to z scores in vec
# normalize the data in vec
scale(vec)
# normalize a single y value relative to dataset x (review the formula for a z score)
y <- mean(vec) - 13
(y-mean(vec)) / sd(vec)
```

Given a sample, you want to know if the mean of the population could reasonably be a particular value m
```{r 9.8 Sampling the Mean of a Sample (t-Test)}
sample <- rnorm(75, mean = 100, sd = 15)
m <- 95

# does a t-test support the population mean of the sample to be 95
t.test(sample, mu = m)
# does a t-test support the population mean of the sample to be 100
t.test(sample, mu = 100)
# what is the default mu argument in the t.test() function
```

Given a sample, you want to determine a confidence interval for the population’s mean.
```{r 9.9 Forming a Confidence Interval for a Mean}
# give a 95% confidence interval for the means of the sample
t.test(sample)
# give a 99% confidence interval for the means of the sample
t.test(sample, conf.level = .99)
```

Given a data sample, what is the confidence interval for the median
```{r9.10 Forming a Confidence Interval for a Median}
# find a 95% conficence interval for the median of sample
wilcox.test(sample, conf.int = TRUE)
# find a 99% confidence interval for the median of sample
wilcox.test(sample, conf.int = 0.99)

```

Given a sample of values from a population consisting of successes and failures, how can you test if the true proportion of successes is p
```{r 9.11 Testing a Sample Proportion}
# what function would be used

# what arguments does the function take

# what does a small p value (typically p<0.05) indicate

# A sports team won 11/20 games last season. How confident should you be that next season the team will win over half their games
```

Given a sample of values from a population consisting of successes and failures. based on the sample data, create a confidence interval for the population's proportion of success
```{r 9.12 Forming a Confidence Interval for a Proportion}
# The price of a good rose 6 times in the last 9 'market events'. Obtain a 99% confidence interval of the true proportion of times the good's price increases in a 'market event'.
prop.test(x = 6, n =9, conf.level = .99)
```

test if each data sample is from a normally distributed population
```{r 9.13 Testing for Normality}
set.seed(100)
sample1 <- sample(rnorm(100), 30)
sample2 <- sample(rexp(100), 30)

# what function can be used to test this
# what arguments does it take
# what is the output and how should it be interpreted

shapiro.test(sample1)
shapiro.test(sample2)

# what is the package "nortest"
# name two graphical tools used to assess normality
```

given a sequence of binary values: yes/no, 0/1, true/false, or other two-valued data, evaluate if the sequence is random?
```{r 9.14 Testing for Runs}
seq1 <- rep(c(TRUE, FALSE), 50)
seq2 <- sample(seq1, 30, replace = TRUE)

tseries::runs.test(as.factor(seq1))
tseries::runs.test(as.factor(seq2))

# what test can be used
# what are the inputs
# what is the output and how should it be interpreted

```

```{r 9.15 Comparing the Means of Two Samples}

```

```{r 9.16 Comparing the Locations of Two Samples Nonparametrically}

```

```{r 9.17 Testing a Correlation for Significance}

```
```{r 9.18 Testing Groups for Equal Proportions}

```
```{r 9.19 Performing Pairwise Comparisons Between Group Means}

```
```{r 9.20 Testing Two Samples for the Same Distribution}

```

