---
title: "Chapter 8"
author: "Colby S."
date: "4/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

calculate the number of combinations of n items taken k at a time.
```{r}
# How many ways can we select 3 items from 5 items?
# How many ways can we select 3 items from 50 items?
# How many ways can we select 30 items from 50 items?

# /hidden choose()
```

generate all combinations of n items taken k at a time.
```{r}

# 
```

generate random numbers
```{r}

```

generate reproducible random numbers
```{r}

```

sample a dataset randomly
```{r}

```

generate a random sequence
```{r}

```

generate a random permutation of a vector
```{r}

```

calculate either the simple or the cumulative probability associated with a discrete random variable
```{r}

```

calculate the distribution function (DF) or cumulative distribution function (CDF) for a continuous random variable
```{r}

```

Given a probability p and a distribution, determine the corresponding quantile for p: the value x such that P(X ≤ x) = p.
```{r}

```

plot the density function of a probability distribution
```{r}

```

