---
title: "Chapter 8"
author: "Colby S."
date: "4/7/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
# would be better without having to load tidyverse, but is convenient
```

calculate the number of combinations of n items taken k at a time.
```{r 8.1 Counting the Number of Combinations}
# How many ways can we select 3 items from 5 items?
# 10

# How many ways can we select 3 items from 50 items?
# 19600

# How many ways can we select 30 items from 50 items?
# 4.712921e+13
```

generate all combinations of n items taken k at a time.
```{r 8.2 Generating Combinations}
# generate all combinations of the numbers 1 through 5 taken three at a time
#      [,1] [,2] [,3] [,4] [,5] [,6] [,7] [,8] [,9] [,10]
# [1,]    1    1    1    1    1    1    2    2    2     3
# [2,]    2    2    2    3    3    4    3    3    4     4
# [3,]    3    4    5    4    5    5    4    5    5     5

# generate all combinations of five treatments (T1, T2, T3, T4, T5) taken three at a time
treatments <- c("T1", "T2", "T3", "T4", "T5")
#      [,1] [,2] [,3] [,4] [,5] [,6] [,7] [,8] [,9] [,10]
# [1,] "T1" "T1" "T1" "T1" "T1" "T1" "T2" "T2" "T2" "T3" 
# [2,] "T2" "T2" "T2" "T3" "T3" "T4" "T3" "T3" "T4" "T4" 
# [3,] "T3" "T4" "T5" "T4" "T5" "T5" "T4" "T5" "T5" "T5" 
```

generate random numbers
```{r 8.3 Generating random numbers}
# generate a random uniform number between 0 and 1

# generate 10 random uniform numbers between 0 and 1

# generate one uniform variate between 1 and 10

# generate one normal variate 

# generate one binomial varaiate

# one poisson variate

# one exponential varaite

# one gamma variate

# generate three normal random values drawn from distributions with means of –10, 0, and +10 with sd of 1

# calculate 30 draws of a normal variate whose mean is itself randomly distributed and with hyperparameters of μ = 0 and σ = 0.2

```

generate reproducible random numbers
```{r 8.4 Generating Reproducible Random Numbers}
# generate two sets of identical numbers using the runif() function
# does the set.seed() function affect the all random generation in a document?
set.seed(100)
runif(10)
runif(10)

set.seed(100)
runif(10)
set.seed(100)
runif(10)
```

sample a dataset randomly
```{r 8.5 Generating a Random Sample}
# take 50 samples(without replacement) of diamond prices from the diamonds dataset
  # the diamonds dataset is loaded in the ggplot package

# take another 50 samples but include replacement
# what does sample() do when size is not specified?


set.seed(42)
x <- rnorm(1000, 4, 10)
medians <- numeric(1000)   # empty vector of 1000 numbers
# sample x 1000 times with replacement and store the median of each sample in the vaiable medians. Use a for loop to do this

```

generate a random sequence
```{r Generating Random Sequences}
# simulate the results of ten coin tosses
sample(c(T,F), 10, replace = T)
# stimulate the results of ten coin tosses of an unfair coin (heads!= 0.5, tails != 0.5)
# what happens when the probabilities entered don't add up to one, how would r convert prob = c(10,20) for the coin flip example?
# generate 10 binomial variates

```

generate a random permutation of a vector
```{r 8.7 Randomly Permuting a Vector}
# generate a random permutation of vector 1:10

```

calculate either the simple or the cumulative probability associated with a discrete random variable
```{r 8.8 Calculating Probabilities for Discrete Distributions}
# if the probability of success is 0.5, what is the probability of getting exactly seven success' in ten trials
# 0.117

## cumulative probability
# what is the probability of having at least 7 success'
# 0.945

## survival probability
# What is the probability of having at most 7 success'
# 0.0546

## interval probability
# what is the probability of having more than 3 but no more than 7 success'
# 0.773

```

calculate the distribution function (DF) or cumulative distribution function (CDF) for a continuous random variable
```{r 8.9 Calculating Probabilites for Continuous Distributions}
# what’s the probability of a draw being below .8 for a draw from a random standard normal distribution?
# 0.788

# calculate the probability that a man is shorter than 66 inches, assuming that men’s heights are normally distributed with a mean of 70 inches and a standard deviation of 3 inches
# 0.0912

# calculate the probability that a man is TALLER than 66 inches, assuming that men’s heights are normally distributed with a mean of 70 inches and a standard deviation of 3 inches
# 0.908

# calculate the probability that an exponential variable with a mean of 40 could be less than 20
# 0.393

# what is the probability that the same exponential variable could be greater than 50

# For the same exponential variable, what is P(20 < X < 50), the probability that it could fall between 20 and 50
# 0.320

```

Given a probability p and a distribution, determine the corresponding quantile for p: the value x such that P(X ≤ x) = p.
```{r 8.10 Converting Probabilites to Quantiles}
# convert the 5th percentile of distribution X~N(100, 15) to a quantile
# 75.327

# calculate the 95% confidence interval of a standard normal variable.
cat("the 95% confidence interavl of a standard normal variable is from", , "to", )
```

plot the density function of a probability distribution
this is really a lesson on ggplot more than it is about density functions
```{r 8.11 Plotting a Density Function}
# library(ggplot)
x <- seq(-3,3,0.1)
dens <- data.frame(x = x,
                   y = dnorm(x))

ggplot(dens, aes(x,y)) +
  geom_line()

# shade a the region between the 75th and 95th percentile. 
x <- seq(-3,3, length.out = 100)
df <- data.frame(x = x,
                   y = dnorm(x, mean = 0, sd = 1))

q75 <- quantile(df$x, 0.75)
q95 <- quantile(df$x, 0.95)

ggplot(dens, aes(x,y)) +
  geom_line() +
  labs(
    title = "Standard Normal Distriution",
    y = "Density",
    x = "Quantile"
  ) +
  geom_ribbon(
    data = subset(df, x > q75 & x < q95),
    aes(ymax = y),
    ymin = 0,
    fill = "blue",
    alpha = 0.5
  )
  
```

